\section{Appendix}
\subsection{Using OpenCLIP}
\subsubsection*{What Are OpenCLIP Models? Are They LLMs?}

OpenCLIP models are \textbf{vision-language encoders} trained to align images and text in a shared embedding space using contrastive-style objectives. In practice, an image encoder (for example ViT or ConvNeXt) and a text encoder are trained so matching image-text pairs have high similarity and non-matching pairs have low similarity. This makes OpenCLIP strong for zero-shot image classification, retrieval, and feature extraction.

OpenCLIP models are \textbf{not LLMs}. LLMs are autoregressive language generators trained to predict the next token and produce long-form text. By contrast, OpenCLIP is primarily an alignment model that returns embeddings/similarity scores rather than free-form language generation.

In short: OpenCLIP answers ``which label/text best matches this image?'' while LLMs answer ``what text should come next?'' Some modern systems combine both, where OpenCLIP-like encoders provide visual grounding and an LLM provides generative reasoning and explanation.

\subsubsection{Step 1: Benchmarking on the Practical Ground Truth Set}
In Step 1, we benchmarked several OpenCLIP-compatible vision-language models on the same \texttt{PracticalGroundTruth500} setup (500 images, 10 classes). For each model, we report:
\begin{itemize}
	\item zero-shot accuracy (text-prompt classification),
	\item few-shot accuracy (prototype-based adaptation using class support examples),
	\item model family,
	\item pretraining dataset/checkpoint tag,
	\item and the main training objective (loss function).
\end{itemize}

\subsubsection*{Model Information}
\begin{enumerate}
    \item \texttt{convnext\_xxlarge} 
        \begin{itemize}
            \item Pretrained tag: \texttt{laion2b\_s34b\_b82k\_augreg\_soup} (LAION-2B).
            \item Family: OpenCLIP ConvNeXt-XXL.
            \item Main objective: CLIP contrastive loss (symmetric image-text cross-entropy).
        \end{itemize}
    \item \texttt{MobileCLIP-S2} 
        \begin{itemize}
            \item Pretrained tag: \texttt{datacompdr\_s12m\_b5m} (DataCompDR).
            \item Family: MobileCLIP.
            \item Main objective: CLIP-style contrastive training within the MobileCLIP reinforced/distillation pipeline.
        \end{itemize}
    \item \textbf{\texttt{coca\_ViT-L-14}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{mscoco\_finetuned\_laion2b\_s13b\_b90k} (LAION-2B with MSCOCO fine-tuning).
            \item Family: CoCa (Contrastive Captioner).
            \item Main objective: weighted contrastive loss plus captioning cross-entropy.
        \end{itemize}

    \item \textbf{\texttt{ViT-bigG-14}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{laion2b\_s39b\_b160k} (LAION-2B).
            \item Family: OpenCLIP ViT-bigG.
            \item Main objective: CLIP contrastive loss.
        \end{itemize}

    \item \textbf{\texttt{ViT-SO400M-16-SigLIP2-384}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{webli} (WebLI).
            \item Family: SigLIP2.
            \item Main objective: pairwise sigmoid loss (SigLIP objective).
        \end{itemize}

    \item \textbf{\texttt{ViT-g-14}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{laion2b\_s34b\_b88k} (LAION-2B).
            \item Family: OpenCLIP ViT-g.
            \item Main objective: CLIP contrastive loss.
        \end{itemize}

    \item \textbf{\texttt{ViT-H-14}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{laion2b\_s32b\_b79k} (LAION-2B).
            \item Family: OpenCLIP ViT-H.
            \item Main objective: CLIP contrastive loss.
        \end{itemize}

    \item \textbf{\texttt{ViT-H-14-CLIPA-336}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{datacomp\_1b\_s13b\_b145k} (DataComp-1B).
            \item Family: CLIPA / OpenCLIP.
            \item Main objective: CLIP-style contrastive loss.
        \end{itemize}

    \item \textbf{\texttt{ViT-H-14-quickgelu}.}
        \begin{itemize}
            \item Pretrained tag: \texttt{dfn5b} (DFN-5B).
            \item Family: DFN / OpenCLIP QuickGELU variant.
            \item Main objective: CLIP contrastive loss.
        \end{itemize}
\end{enumerate}

\subsubsection*{Accuracy Table (500-Image Practical Set)}
\begin{table}[H]
\centering
\caption{Pipeline 4 accuracy comparison on \texttt{PracticalGroundTruth500} (500 images)}
\label{tab:part3_pipeline4_zerofewshot}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.58\textwidth}cc}
\toprule
\textbf{Model} & \textbf{Zero-shot (\%)} & \textbf{Few-shot (\%)} \\
\midrule
\texttt{convnext\_xxlarge} & 45.2 & 91.4 \\
\texttt{MobileCLIP-S2} & 54.4 & 94.2 \\
\texttt{coca\_ViT-L-14} & 46.2 & 87.4 \\
\texttt{ViT-bigG-14} & 16.03 & 93.30 \\
\texttt{ViT-SO400M-16-SigLIP2-384} & 7.4 & 93.40 \\
\texttt{ViT-g-14} & 16.05 & 93.45 \\
\texttt{ViT-H-14} & 16.03 & 93.10 \\
\texttt{ViT-H-14-CLIPA-336} & 16.03 & 93.21 \\
\texttt{ViT-H-14-quickgelu} & 16.35 & 93.07 \\
\bottomrule
\end{tabular}
\end{table}
