\subsection{Problem 2: Convolutional Neural Network (CNN) on ReducedMNIST}

This problem trains CNN classifiers directly from the raw $28 \times 28$ ReducedMNIST images, without a separate feature extraction stage. The starting model is a LeNet-style network adjusted for $28 \times 28$ grayscale inputs and uses ReLU activations, as required by the assignment.

\subsubsection{CNN Architecture and Variations}
The baseline model contains two $5 \times 5$ convolutional layers, max-pooling after each convolution, and three fully connected layers for classification. Four CNN entries are reported in the assignment table:
\begin{itemize}
    \item \textbf{Variation 1: Baseline LeNet with ReLU} using 6 and 16 convolutional filters.
    \item \textbf{Variation 2: Wider network} with 12 and 32 convolutional filters and wider fully connected layers.
    \item \textbf{Variation 3: LeakyReLU network} replacing ReLU by LeakyReLU.
    \item \textbf{Variation 4: Dropout network} adding dropout regularization before the later fully connected layers.
\end{itemize}

\subsubsection{Comparison with Assignment 1 and Problem 1}
Table~\ref{tab:prob2_results} follows the assignment comparison format. Assignment 1 processing time is reported as total feature extraction, training, and testing time. MLP processing time is reported as training plus testing time from Problem 1. CNN time is split into training and testing time because no external feature extraction step is used. All times are in milliseconds (msec) and all accuracies are rounded to one decimal place.

\begin{table}[H]
    \centering
    \caption{Performance comparison across Assignment 1 classical methods, Problem 1 MLP, and Problem 2 CNN}
    \label{tab:prob2_results}
    \resizebox{\linewidth}{!}{
    \begin{tabular}{|l|c|c|c|c|c|c|}
        \hline
        \multirow{3}{*}{\textbf{Classifier / Variation}} & \multicolumn{6}{c|}{\textbf{Feature Representation}} \\
        \cline{2-7}
        & \multicolumn{2}{c|}{\textbf{DCT}} & \multicolumn{2}{c|}{\textbf{PCA}} & \multicolumn{2}{c|}{\textbf{AutoEncoder}} \\
        \cline{2-7}
        & \textbf{Accuracy} & \textbf{Processing Time (msec)} & \textbf{Accuracy} & \textbf{Processing Time (msec)} & \textbf{Accuracy} & \textbf{Processing Time (msec)} \\
        \hline
        \multicolumn{7}{|l|}{\textbf{Assignment 1: DCT/PCA feature classifiers}} \\
        \hline
        K-means, $K=1$  & 78.1\% & 4800.0 & 85.2\% & 2464.6 & 22.3\% & 1721.5 \\
        K-means, $K=4$  & 84.4\% & 5077.3 & 82.6\% & 3608.5 & 24.0\% & 627.7 \\
        K-means, $K=16$ & 88.8\% & 12072.5 & 85.7\% & 5800.8 & 41.1\% & 827.8 \\
        K-means, $K=32$ & 89.7\% & 38344.4 & 85.6\% & 8701.1 & 43.8\% & 1029.3 \\
        SVM, linear & 90.3\% & 11487.5 & 90.7\% & 10626.8 & 27.9\% & 2195.7 \\
        SVM, nonlinear RBF, $\gamma=\mathrm{scale}$ & 95.6\% & 20560.2 & 95.2\% & 21054.4 & 23.2\% & 3651.0 \\
        \hline
        \multicolumn{7}{|l|}{\textbf{Assignment 2 Problem 1: MLP with DCT features}} \\
        \hline
        MLP, 1-hidden layer & 97.3\% & 7225.5 & -- & -- & 10.0\% & 6131.1 \\
        MLP, 3-hidden layers & 97.3\% & 10114.9 & -- & -- & 28.0\% & 11635.5 \\
        MLP, 4-hidden layers & 97.3\% & 11957.3 & -- & -- & 10.0\% & 12462.8 \\
        \hline
        \multicolumn{7}{|c|}{\textbf{Assignment 2 Problem 2: CNN directly from images, no feature extraction}} \\
        \hline
        \textbf{CNN variation} & \textbf{Accuracy} & \multicolumn{2}{c|}{\textbf{Training Time (msec)}} & \multicolumn{3}{c|}{\textbf{Testing Time (msec)}} \\
        \hline
        Variation 1: Baseline LeNet with ReLU & 94.9\% & \multicolumn{2}{c|}{90467.0} & \multicolumn{3}{c|}{14418.7} \\
        Variation 2: Wider filters/FC layers & 97.0\% & \multicolumn{2}{c|}{19062.4} & \multicolumn{3}{c|}{503.3} \\
        Variation 3: LeakyReLU activation & 96.3\% & \multicolumn{2}{c|}{15780.3} & \multicolumn{3}{c|}{433.4} \\
        Variation 4: Dropout regularization & 97.0\% & \multicolumn{2}{c|}{16025.5} & \multicolumn{3}{c|}{426.3} \\
        \hline
    \end{tabular}
    }
\end{table}

\subsubsection{Discussion}
The wider CNN and dropout CNN reached the strongest CNN accuracy at $97.0\%$, close to the DCT-based MLP result of $97.3\%$. The baseline CNN was slower in the recorded run and lower in accuracy, while the modified CNNs reduced testing time substantially. Compared with the Assignment 1 DCT/PCA classifiers, the neural-network models give higher accuracy on the same ReducedMNIST task. The AutoEncoder feature columns show significantly lower accuracies (ranging from 10\% to 43.8\%) compared to DCT and PCA, which suggests that the simple AutoEncoder bottleneck representation requires further tuning (e.g., more epochs or a different architecture) to match the classical feature extractors on the ReducedMNIST dataset.
